---
title: "Mod 1 & 2 (Wk 1)"
author: "Julianna D."
date: "2023-09-07"
output: html_document
---
# Module 1: First Steps with R

## Challenge 1: 
```{r}
# Basic Math
8+5
10-6/2
(10-6)/2
10*5
15/5
10^5
3*pi
```

```{r}
# Assignments
x<-6
x
y<-5
y
z<-x*y
z
x2<-x^2
x2
```

```{r}
# Built-in Functions
x<-10
log(x)
factorial(x)
y<-81
sqrt(y)
z<--8.349218
round(z, digits=3)
abs(z*y)
```

```{r}
# Use the `ls()` function to list the variables currently stored in your active session. How many do you have?
ls()
```

```{r}
# Clear all the variables you've defined
rm(list=ls())
```

What happens if you type a function name without including the parentheses?
Answer: if you type a function without including the parentheses, the output will show you the details of the function, such as the necessary arguments and what each means.
```{r}
ls
```

## Challenge 2: Objects, Classes, Attributes, Vectors
```{r}
#using c() to create vectors
x <- c(15, 16, 12, 3, 21, 45, 23)
x
y <- c("once", "upon", "a", "time")
y
z <- "once upon a time"
z
```

```{r}
f <- function(){
	# code to be evaluated
}
# this is the minimal definition for a function
```

- What is the class of x? Of y? of f? Use the `class()` function to check.
```{r}
class(x)
class(y)
class(f)
```

What is the class of X now?
Answer: it is now a character.
```{r}
x <- c("2", 2, "zombies")
class(x)
```

