---
title: "Mod 18 (Wk 11)"
author: "Julianna D."
date: "2023-11-16"
output: html_document
---

# Module 18 Mixed Effects Modeling

```{r, message=FALSE}
library(curl)
library(ggplot2)
library(lme4)
library(AICcmodavg)
library(MuMIn)
```

Varieties of Mixed Models
- Linear Mixed Models (LMM)

- Generalized Linear Mixed Models (GLMM)

- Nonlinear Mixed Models (NLMM)

Mixed models include BOTH fixed & random effects.

## Challenge 1
```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall21/chimpgrooming.csv")
d <- read.csv(f, header = TRUE, sep = ",")
head(d)
summary(d)
```

Visualization of grooming received in relation to subject ID

```{r}
par(mfrow = c(1, 1))
boxplot(data = d, duration ~ subject, col = c("lightpink1"))
```

Grooming Received in relation to reprod. cond.

```{r}
boxplot(data = d, duration ~ reprocondition, col = c("burlywood2", "lightpink1"))
```

Grooming received in relation to reprod. cond. and parity

```{r}
boxplot(data = d, duration ~ reprocondition * parity, col = c("burlywood2",
    "lightpink1"))
```

```{r}
boxplot(data = d, duration ~ reprocondition * subject, col = c("burlywood2",
    "lightpink1"))
```
### Random Intercept Models

```{r}
lme <- lmer(data = d, duration ~ reprocondition + parity + (1 | subject))
summary(lme)
```

```{r}
coefficients(lme)
```
Statistical Significance

```{r}
full <- lmer(data = d, duration ~ reprocondition + parity + (1 | subject), REML = FALSE)
summary(full)
```

```{r}
reduced <- lmer(data = d, duration ~ parity + (1 | subject), REML = FALSE)
summary(reduced)
```

Likelihood Ratio Test using anova()
```{r}
anova(reduced, full, test = "Chisq")
```

```{r}
full <- lmer(data = d, duration ~ reprocondition + parity + (1 | subject), REML = FALSE)
reduced <- lmer(data = d, duration ~ reprocondition + (1 | subject), REML = FALSE)
anova(reduced, full, test = "Chisq")
```
Including parity singificantly improves the fit of our model!


## Challenge 2

```{r}
full <- lmer(data = d, duration ~ reprocondition * parity + (1 | subject), REML = FALSE)
reduced <- lmer(data = d, duration ~ reprocondition + parity + (1 | subject),
    REML = FALSE)
anova(reduced, full, test = "Chisq")
```
### Random Slope Models

```{r}
lme <- lmer(data = d, duration ~ reprocondition + parity + (1 + reprocondition |
    subject) + (1 + parity | subject), REML = FALSE)
summary(lme)
```

```{r}
coefficients(lme)
```

Reproductive Condition
```{r}

full <- lmer(data = d, duration ~ reprocondition + parity + (1 + reprocondition |
    subject) + (1 + parity | subject), REML = FALSE)

reduced <- lmer(data = d, duration ~ parity + (1 + reprocondition | subject) +
    (1 + parity | subject), REML = FALSE)
```
```{r}
anova(reduced, full, test = "Chisq")
```
Parity
```{r}
full <- lmer(data = d, duration ~ reprocondition + parity + (1 + reprocondition |
    subject) + (1 + parity | subject), REML = FALSE)

null <- lmer(data = d, duration ~ reprocondition + (1 + reprocondition | subject) +
    (1 + parity | subject), REML = FALSE)
```

```{r}
anova(reduced, full, test = "Chisq")
```

Determining Model Fit

## Challenge 3

```{r}
print(aictab(list(full, reduced, null), c("full", "reduced", "null")), LL = FALSE)
```

```{r}
r.squaredGLMM(full)
```

```{r}
r.squaredGLMM(reduced)
```

```{r}
r.squaredGLMM(null)
```

