---
title: "Mod11-12 (Wk 6)"
author: "Julianna D."
date: "2023-10-12"
output: html_document
---

# Module 11 Type I and Type II Error and Power

Type 1 Error & Multiple Testing Problem
```{r}
typeI <- function(mu0, sigma, n, alternative = "two.tailed", alpha = 0.05, k = 10000) {
}
```

```{r}
typeI <- function(mu0, sigma, n, alternative = "two.tailed", alpha = 0.05, k = 1000) {
    p <- rep(NA, k)  # sets up a vector of empty p values
    for (i in 1:k) {
        # sets up a loop to run k simulations
        x <- rnorm(n = n, mean = mu0, sd = sigma)  # draws a sample from our distribution
        m <- mean(x)  # calculates the mean
        s <- sd(x)  # calculates the standard deviation
        z <- (m - mu0)/(s/sqrt(n))  # calculates the T statistic for the sample drawn from the null distribution relative to the null distribution
        # alternatively use t <- (m-mu0)/(s/sqrt(n))
        if (alternative == "less") {
            p[i] <- pnorm(z, lower.tail = TRUE)  # calculates the associated p value
            # alternatively, use p[i] <- pt(t,df=n-1,lower.tail=TRUE)
        }
        if (alternative == "greater") {
            p[i] <- pnorm(z, lower.tail = FALSE)  # calculates the associated p value
            # alternatively, use p[i] <- pt(t,df=n-1,lower.tail=FALSE)
        }
        if (alternative == "two.tailed") {
            if (z > 0)
                {
                  p[i] <- 2 * pnorm(z, lower.tail = FALSE)
                }  # alternatively, use if (t > 0) {p[i] <- pt(t,df=n-1,lower.tail=FALSE)}
            if (z < 0)
                {
                  p[i] <- 2 * pnorm(z, lower.tail = TRUE)
                }  # alternatively, use if (t < 0) {p[i] <- pt(t,df=n-1,lower.tail=TRUE)}
        }
    }

    curve(dnorm(x, mu0, sigma/sqrt(n)), mu0 - 4 * sigma/sqrt(n), mu0 + 4 * sigma/sqrt(n),
        main = paste("Sampling Distribution Under the Null Hypothesis\nType I error rate from simulation = ",
            length(p[p < alpha])/k, sep = ""), xlab = "x", ylab = "Pr(x)", col = "red",
        xlim = c(mu0 - 4 * sigma/sqrt(n), mu0 + 4 * sigma/sqrt(n)), ylim = c(0,
            dnorm(mu0, mu0, sigma/sqrt(n))))
    abline(h = 0)

    if (alternative == "less") {
        polygon(cbind(c(mu0 - 4 * sigma/sqrt(n), seq(from = mu0 - 4 * sigma/sqrt(n),
            to = mu0 - qnorm(1 - alpha) * sigma/sqrt(n), length.out = 100),
            mu0 - qnorm(1 - alpha) * sigma/sqrt(n))), c(0, dnorm(seq(from = mu0 -
            4 * sigma/sqrt(n), to = mu0 - qnorm(1 - alpha) * sigma/sqrt(n),
            length.out = 100), mean = mu0, sd = sigma/sqrt(n)), 0), border = "black",
            col = "grey")
        q <- pnorm(mu0 - qnorm(1 - alpha) * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) -
            pnorm(mu0 - 4 * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n))
    }
    if (alternative == "greater") {
        polygon(cbind(c(mu0 + qnorm(1 - alpha) * sigma/sqrt(n), seq(from = mu0 +
            qnorm(1 - alpha) * sigma/sqrt(n), to = mu0 + 4 * sigma/sqrt(n),
            length.out = 100), mu0 + 4 * sigma/sqrt(n))), c(0, dnorm(seq(from = mu0 +
            qnorm(1 - alpha) * sigma/sqrt(n), to = mu0 + 4 * sigma/sqrt(n),
            length.out = 100), mean = mu0, sd = sigma/sqrt(n)), 0), border = "black",
            col = "grey")
        q <- pnorm(mu0 + 4 * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) -
            pnorm(mu0 + qnorm(1 - alpha) * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n))
    }
    if (alternative == "two.tailed") {
        polygon(cbind(c(mu0 - 4 * sigma/sqrt(n), seq(from = mu0 - 4 * sigma/sqrt(n),
            to = mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), length.out = 100),
            mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n))), c(0, dnorm(seq(from = mu0 -
            4 * sigma/sqrt(n), to = mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n),
            length.out = 100), mean = mu0, sd = sigma/sqrt(n)), 0), border = "black",
            col = "grey")
        polygon(cbind(c(mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), seq(from = mu0 +
            qnorm(1 - alpha/2) * sigma/sqrt(n), to = mu0 + 4 * sigma/sqrt(n),
            length.out = 100), mu0 + 4 * sigma/sqrt(n))), c(0, dnorm(seq(from = mu0 +
            qnorm(1 - alpha/2) * sigma/sqrt(n), to = mu0 + 4 * sigma/sqrt(n),
            length.out = 100), mean = mu0, sd = sigma/sqrt(n)), 0), border = "black",
            col = "grey")
        q <- pnorm(mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) -
            pnorm(mu0 - 4 * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) +
            pnorm(mu0 + 4 * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n)) -
            pnorm(mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), mean = mu0, sd = sigma/sqrt(n))
    }
    # print(round(q,digits=3)) # this prints area in the shaded portion(s)
    # of the curve
    return(length(p[p < alpha])/k)  # returns the proportion of simulations where p < alpha
}
```

```{r}
eI <- typeI(mu0 = -3, sigma = 2, n = 5000, alternative = "greater", alpha = 0.05)
```

```{r}
eI <- typeI(mu0 = 5, sigma = 2, n = 1000, alternative = "less", alpha = 0.01)
```

## Challenge 1
Multiple Comparison Corrections

```{r}
alpha <- 0.05
pvals <- c(1e-04, 0.003, 0.005, 0.01, 0.02, 0.04, 0.045, 0.11, 0.18, 0.23)
sig <- pvals <= alpha/length(pvals)
sig
```

```{r}
library(ggplot2)
alpha <- 0.05
psig <- NULL
pvals <- c(1e-04, 0.003, 0.005, 0.01, 0.02, 0.04, 0.045, 0.11, 0.18, 0.27)
for (i in 1:length(pvals)) {
    psig[i] <- alpha * i/length(pvals)
}
d <- data.frame(cbind(rank = c(1:10), pvals, psig))
p <- ggplot(data = d, aes(x = rank, y = pvals)) + geom_point() + geom_line(aes(x = rank,
    y = psig))
p
```

```{r}
sig <- pvals <= psig
sig
```
```{r}
sig <- p.adjust(pvals, method = "bonferroni") <= 0.05
sig
```
```{r}
sig <- p.adjust(pvals, method = "BH") <= 0.05
sig
```

Type II Error
```{r}
typeII <- function(mu0, muA, sigma, n, alternative = "two.tailed", alpha = 0.05,
    k = 1000) {
    p <- rep(NA, k)  # sets up a vector of empty p values
    for (i in 1:k) {
        x <- rnorm(n = n, mean = muA, sd = sigma)  # draw from Ha
        m <- mean(x)
        s <- sd(x)
        z <- (m - mu0)/(s/sqrt(n))  # calculates the Z statistic for the sample drawn from Ha relative to the null distribution
        if (alternative == "less") {
            p[i] <- pnorm(z, lower.tail = TRUE)  # calculates the associated p value
            hyp <- "muA < mu0"
        }
        if (alternative == "greater") {
            p[i] <- pnorm(z, lower.tail = FALSE)
            hyp <- "muA > mu0"
        }
        if (alternative == "two.tailed") {
            if (z > 0) {
                p[i] <- 2 * pnorm(z, lower.tail = FALSE)
            }
            if (z < 0) {
                p[i] <- 2 * pnorm(z, lower.tail = TRUE)
            }
            hyp <- "muA ≠ mu0"
        }
    }

    curve(dnorm(x, mu0, sigma/sqrt(n)), mu0 - 4 * sigma/sqrt(n), mu0 + 4 * sigma/sqrt(n),
        main = paste("Sampling Distributions Under the Null (red)\nand Alternative Hypotheses (blue)\nType II error rate from simulation = ",
            length(p[p >= alpha])/k, sep = ""), xlab = "x", ylab = "Pr(x)",
        col = "red", xlim = c(min(c(mu0 - 4 * sigma/sqrt(n), muA - 4 * sigma/sqrt(n))),
            max(c(mu0 + 4 * sigma/sqrt(n), muA + 4 * sigma/sqrt(n)))), ylim = c(0,
            max(c(dnorm(mu0, mu0, sigma/sqrt(n))), dnorm(muA, muA, sigma/sqrt(n)))))

    curve(dnorm(x, muA, sigma/sqrt(n)), muA - 4 * sigma/sqrt(n), muA + 4 * sigma/sqrt(n),
        col = "blue", add = TRUE)
    abline(h = 0)

    if (alternative == "less") {
        polygon(cbind(c(mu0 - qnorm(1 - alpha) * sigma/sqrt(n), seq(from = mu0 -
            qnorm(1 - alpha) * sigma/sqrt(n), to = muA + 4 * sigma/sqrt(n),
            length.out = 100), muA + 4 * sigma/sqrt(n))), c(0, dnorm(seq(mu0 -
            qnorm(1 - alpha) * sigma/sqrt(n), to = muA + 4 * sigma/sqrt(n),
            length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0), border = "black",
            col = "grey")
        abline(v = mu0 - qnorm(1 - alpha) * sigma/sqrt(n), col = "black", lty = 3,
            lwd = 2)
    }

    if (alternative == "greater") {
        polygon(cbind(c(muA - 4 * sigma/sqrt(n), seq(from = muA - 4 * sigma/sqrt(n),
            to = mu0 + qnorm(1 - alpha) * sigma/sqrt(n), length.out = 100),
            mu0 + qnorm(1 - alpha) * sigma/sqrt(n))), c(0, dnorm(seq(from = muA -
            4 * sigma/sqrt(n), to = mu0 + qnorm(1 - alpha) * sigma/sqrt(n),
            length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0), border = "black",
            col = "grey")
        abline(v = mu0 + qnorm(1 - alpha) * sigma/sqrt(n), col = "black", lty = 3,
            lwd = 2)
    }

    if (alternative == "two.tailed") {
        abline(v = mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), col = "black",
            lty = 3, lwd = 2)
        abline(v = mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), col = "black",
            lty = 3, lwd = 2)

        if (z > 0) {
            # greater
            polygon(cbind(c(muA - 4 * sigma/sqrt(n), seq(from = muA - 4 * sigma/sqrt(n),
                to = mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n), length.out = 100),
                mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n))), c(0, dnorm(seq(from = muA -
                4 * sigma/sqrt(n), to = mu0 + qnorm(1 - alpha/2) * sigma/sqrt(n),
                length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0), border = "black",
                col = "grey")
        }

        # less
        if (z < 0) {
            polygon(cbind(c(mu0 - qnorm(1 - alpha/2) * sigma/sqrt(n), seq(from = mu0 -
                qnorm(1 - alpha/2) * sigma/sqrt(n), to = muA + 4 * sigma/sqrt(n),
                length.out = 100), muA + 4 * sigma/sqrt(n))), c(0, dnorm(seq(mu0 -
                qnorm(1 - alpha/2) * sigma/sqrt(n), to = muA + 4 * sigma/sqrt(n),
                length.out = 100), mean = muA, sd = sigma/sqrt(n)), 0), border = "black",
                col = "grey")
        }
    }

    return(length(p[p >= alpha])/k)
}
```

## Challenge 2
```{r}
eII <- typeII(mu0 = 2, muA = 4, sigma = 3, n = 6, alternative = "greater")
```
```{r}
eII <- typeII(mu0 = 5, muA = 2, sigma = 4, n = 18, alternative = "less")
```

```{r}
eII <- typeII(mu0 = 5, muA = 7, sigma = 2, n = 15, alternative = "two.tailed")
```

Power and Effect Size

## Challenge 3
```{r}
library(ggplot2)
library(manipulate)
power.test <- function(mu0, muA, sigma, alpha = 0.05, type, alternative) {
    p <- 0
    for (i in 2:200) {
        x <- power.t.test(n = i, delta = abs(muA - mu0), sd = sigma, sig.level = alpha,
            power = NULL, type = type, alternative = alternative)
        p <- c(p, x$power)
    }
    d <- data.frame(cbind(1:200, p, 1 - p))
    critn <- 0
    for (i in 1:199) {
        if (p[i] < 0.8 && p[i + 1] >= 0.8) {
            critn <- i + 1
        } else {
            critn <- critn
        }
    }
    names(d) <- c("n", "power", "beta")
    g <- ggplot(data = d) + xlab("sample size n") + ylab("Type II Error Rate, Beta  (Red)\nand\nPower, 1-Beta (Blue)") +
        ggtitle("Power for T Tests\n(assuming equal n and variance across the two groups)") +
        ylim(0, 1) + geom_point(aes(x = n, y = power), colour = "blue", alpha = 1/2) +
        geom_line(aes(x = n, y = power), colour = "blue", alpha = 1/2) + geom_line(aes(x = n,
        y = 0.8), colour = "red", lty = 3) + geom_point(aes(x = n, y = beta),
        colour = "red", alpha = 1/2) + geom_line(aes(x = n, y = beta), colour = "red",
        alpha = 1/2) + geom_linerange(aes(x = critn, ymin = 0, ymax = 0.8),
        colour = "blue", alpha = 1/4) + annotate("text", x = 150, y = 0.5, label = paste("Effect Size = ",
        round(abs(mu0 - muA)/sigma, digits = 3), "\nCritical n = ", critn, sep = ""))
    print(g)
}
```

# Module 12 Intro to Linear Regression

```{r, message=FALSE}
library(curl)
library(ggplot2)
library(gridExtra)
library(manipulate)
library(lmodel2)
```

Covariance & Correlation

```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

```{r}
plot(data = d, height ~ weight)
```

## Challenge 1
```{r}
w <- d$weight
h <- d$height
n <- length(w)  # or length(h)
cov_wh <- sum((w - mean(w)) * (h - mean(h)))/(n - 1)
cov_wh
#same outcome using base R
cov(w, h)
```

## Challenge 2
```{r}
sd_w <- sd(w)
sd_h <- sd(h)
cor_wh <- cov_wh/(sd_w * sd_h)
cor_wh
```

Regression & The lm() function

## Challenge 3
- plot the zombie height as a function of age
```{r}
plot(data = d, height ~ age)
```

```{r}
beta1 <- cor(d$height, d$age) * sd(d$height)/sd(d$age)
beta1
beta0 <- mean(d$height) - beta1 * mean(d$age)
beta0
```

```{r}
m <- lm(height ~ age, data = d)
m
```

Statistical Inference in Regression + Interpreting Regression Coefs & Prediction
```{r}
m <- lm(data = d, height ~ weight)
summary(m)
```

```{r}
t <- coef(summary(m))
t <- data.frame(unlist(t))
colnames(t) <- c("Est", "SE", "t", "p")
t
```
```{r}
t$calct <- (t$Est - 0)/t$SE
t$calcp <- 2 * pt(t$calct, df = 998, lower.tail = FALSE)  # x2 because is 2-tailed test
t
```
```{r}
t$lower <- t$Est - qt(0.975, df = 998) * t$SE
t$upper <- t$Est + qt(0.975, df = 998) * t$SE
ci <- c(t$lower, t$upper)  # by hand
ci
```
```{r}
ci <- confint(m, level = 0.95)  # using the results of lm()
ci
```

## Challenge 4
If zombie survivor weight is measured in pounds and zombie survivor height is measured in inches, what is the expected height of a zombie survivor weighing 150 pounds?

```{r}
beta0 <- t$Est[1]
beta1 <- t$Est[2]
h_hat <- beta1 * 150 + beta0
h_hat
```

```{r}
h_hat_difference <- (beta1 * 220 + beta0) - (beta1 * 180 + beta0)
h_hat_difference
```

```{r}
m <- lm(data = d, height ~ weight)
h_hat <- predict(m, newdata = data.frame(weight = d$weight))
df <- data.frame(cbind(d$weight, d$height, h_hat))
names(df) <- c("x", "y", "yhat")
head(df)
```

```{r}
g <- ggplot(data = df, aes(x = x, y = yhat))
g <- g + geom_point()
g <- g + geom_point(aes(x = x, y = y), colour = "red")
g <- g + geom_segment(aes(x = x, y = yhat, xend = x, yend = y))
g
```
```{r}
ci <- predict(m, newdata = data.frame(weight = 150), interval = "confidence",
    level = 0.95)  # for a single value
ci
```

```{r}
ci <- predict(m, newdata = data.frame(weight = d$weight), interval = "confidence",
    level = 0.95)  # for a vector of values
head(ci)
```
```{r}
df <- cbind(df, ci)
names(df) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr")
head(df)
```

```{r}
g <- ggplot(data = df, aes(x = x, y = y))
g <- g + geom_point(alpha = 1/2)
g <- g + geom_line(aes(x = x, y = CIfit), colour = "black")
g <- g + geom_line(aes(x = x, y = CIlwr), colour = "blue")
g <- g + geom_line(aes(x = x, y = CIupr), colour = "blue")
g
```

```{r}
# for y at each x
pi <- predict(m, newdata = data.frame(weight = 150), interval = "prediction",
    level = 0.95)  # for a single value
pi

#for a vector of values
pi <- predict(m, newdata = data.frame(weight = d$weight), interval = "prediction",
    level = 0.95)  # for a vector of values
head(pi)
```
```{r}
df <- cbind(df, pi)
names(df) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr", "PIfit", "PIlwr",
    "PIupr")
head(df)
```
```{r}
g <- g + geom_line(data = df, aes(x = x, y = PIlwr), colour = "red")
g <- g + geom_line(data = df, aes(x = x, y = PIupr), colour = "red")
g
```

## Challenge 5
Construct a linear model for the regression of zombie survivor height on age and predict the mean height, the 95% confidence interval (CI) around the predicted mean height, and the 95% prediction interval (PI) around that mean for a vector of zombie survivor ages, v <- seq(from=10, to=30, by=1)
```{r}
v <- seq(from = 10, to = 30, by = 1)
m <- lm(data = d, height ~ age)
ci <- predict(m, newdata = data.frame(age = v), interval = "confidence", level = 0.95)
pi <- predict(m, newdata = data.frame(age = v), interval = "prediction", level = 0.95)
```

Then, plot your points, your regression line, and lines for the lower and upper limits of the CI and of the PI.
```{r}
plot(data = d, height ~ age)
lines(x = v, y = ci[, 1], col = "black")
lines(x = v, y = ci[, 2], col = "blue")
lines(x = v, y = ci[, 3], col = "blue")
lines(x = v, y = pi[, 2], col = "red")
lines(x = v, y = pi[, 3], col = "red")
```

```{r}
require(gridExtra)
df <- data.frame(cbind(v, ci, pi))
names(df) <- c("age", "CIfit", "CIlwr", "CIupr", "PIfit", "PIlwr", "PIupr")
head(df)
```
```{r}
g1 <- ggplot(data = d, aes(x = age, y = height))
g1 <- g1 + geom_point(alpha = 1/2)
g1 <- g1 + geom_line(data = df, aes(x = v, y = CIfit), colour = "black", lwd = 1)
g1 <- g1 + geom_line(data = df, aes(x = v, y = CIlwr), colour = "blue")
g1 <- g1 + geom_line(data = df, aes(x = v, y = CIupr), colour = "blue")
g1 <- g1 + geom_line(data = df, aes(x = v, y = PIlwr), colour = "red")
g1 <- g1 + geom_line(data = df, aes(x = v, y = PIupr), colour = "red")
g2 <- ggplot(data = d, aes(x = age, y = height))
g2 <- g2 + geom_point(alpha = 1/2)
g2 <- g2 + geom_smooth(method = "lm", formula = y ~ x)
grid.arrange(g1, g2, ncol = 2)
```

