---
title: "Mod13 (Wk 7)"
author: "Julianna D."
date: "2023-10-19"
output: html_document
---
# Module 13 Elements of Regression Analysis

```{r}
library(curl)
library(car)
```

```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```

Analysis of Variance and ANOVA Tables

```{r}
m <- lm(data = d, height ~ weight)
SSY <- sum((m$model$height - mean(m$model$height))^2)  # height - mean(height)
SSY
```

```{r}
SSR <- sum((m$fitted.values - mean(m$model$height))^2)  # predicted height - mean height
SSR
```

```{r}
SSE <- sum((m$model$height - m$fitted.values)^2)  # height - predicted height
SSE
```

```{r}
df_regression <- 1
df_error <- 998
df_y <- 999
MSR <- SSR/df_regression
MSE <- SSE/df_error
MSY <- SSY/df_y

fratio <- MSR/MSE
fratio
```

Evaluate F ratio test statistic against F distribution
```{r}
curve(df(x, df = 1, df2 = 1), col = "green", lty = 3, lwd = 2, xlim = c(0, 10),
    main = "Some Example F Distributions\n(vertical line shows critical value for df1=1,df2=998)",
    ylab = "f(x)", xlab = "x")
curve(df(x, df = 2, df2 = 2), col = "blue", lty = 3, lwd = 2, add = TRUE)
curve(df(x, df = 4, df2 = 4), col = "red", lty = 3, lwd = 2, add = TRUE)
curve(df(x, df = 8, df2 = 6), col = "purple", lty = 3, lwd = 2, add = TRUE)
curve(df(x, df = 1, df2 = 998), col = "black", lwd = 3, add = TRUE)
legend("top", c("df1=1,df2=1", "df1=2,df2=2", "df1=4,df2=4", "df1=8,df2=6",
    "df1=1,df2=998"), lty = 3, lwd = 2, col = c("green", "blue", "red", "purple",
    "black"), bty = "n", cex = 0.75)

fcrit <- qf(p = 0.95, df1 = 1, df2 = 998)
fcrit
abline(v = fcrit)
abline(h = 0)
polygon(cbind(c(fcrit, seq(from = fcrit, to = 10, length.out = 1000), 8), c(0,
    df(seq(from = fcrit, to = 8, length.out = 1000), df1 = 1, df2 = 998), 0)),
    border = "black", col = "grey")
```

```{r}
1 - pf(q = fratio, df1 = 1, df2 = 998)
```

```{r}
a <- aov(data = d, height ~ weight)
summary(a)

summary.aov(m)

rsquared <- SSR/SSY
rsquared

rho <- sqrt(rsquared)
rho
```

Standard Errors of Regression Coefficients
```{r}
SSX <- sum((m$model$weight - mean(m$model$weight))^2)
SEbeta1 <- sqrt(MSE/SSX)
SEbeta1
```

```{r}
SEbeta0 <- sqrt((MSE * sum(m$model$weight^2))/(1000 * SSX))
SEbeta0
```

```{r}
SEyhat <- sqrt(MSE * (1/1000 + (m$model$weight - mean(m$model$weight))^2/SSX))
head(SEyhat)
```

```{r}
summary(m)
```

## Challenge 1

Calculate the residuals from the regression of zombie height on weight and plot these in relation to weight (the x variable).
```{r}
m <- lm(data = d, height ~ weight)
plot(x = d$weight, y = m$residuals)
# or
e <- resid(m)
plot(x = d$weight, y = e)

```

Now, plot a histogram of your residuals… ideally they are normally distributed!
```{r}
hist(e, xlim = c(-4 * sd(e), 4 * sd(e)), breaks = 20, main = "Histogram of Residuals")
```

```{r}
plot(m$model$weight, m$residuals)
```

```{r}
par(mfrow = c(2, 2))
plot(m)
```

```{r}
qqnorm(m$residuals)
```

```{r}
qqPlot(m$residuals)
```

Tests for normality
```{r}
s <- shapiro.test(m$residuals)
s
```

## Challenge 2

```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

```{r}
plot(data = d, WeaningAge_d ~ Body_mass_female_mean)
```

```{r}
model <- lm(data = d, WeaningAge_d ~ Body_mass_female_mean)
summary(model)
```

```{r}
plot(model)
```

```{r}
qqPlot(model$residuals)
```

```{r}
# Test for normality
s <- shapiro.test(model$residuals)
s
```

Data Transformations

## Challenge 3
```{r}
d$logWeaningAge <- log(d$WeaningAge_d)
d$logFemaleBodyMass <- log(d$Body_mass_female_mean)
plot(data = d, logWeaningAge ~ logFemaleBodyMass)
```

```{r}
model <- lm(data = d, logWeaningAge ~ logFemaleBodyMass)
summary(model)
```

```{r}
qqPlot(model$residuals)
```

```{r}
s <- shapiro.test(model$residuals)
s
```

```{r}
par(mfrow = c(1, 2))

a <- 2
b <- 2

# log x
x <- seq(from = 0, to = 100, length.out = 1000)
y <- a + b * log(x)
plot(x, y, type = "l", main = "untransformed")
plot(log(x), y, type = "l", main = "log(x)")
```

```{r}
# log y
x <- seq(from = 0, to = 10, length.out = 1000)
y <- exp(a + b * x)
plot(x, y, type = "l", main = "untransformed")
plot(x, log(y), type = "l", main = "log(y)")
```

```{r}
# assymptotic
x <- seq(from = 1, to = 100, length.out = 100)
y <- (a * x)/(1 + b * x)
plot(x, y, type = "l", main = "untransformed")
plot(1/x, y, type = "l", main = "1/x")
```

```{r}
# reciprocal
x <- seq(from = 1, to = 100, length.out = 100)
y <- a + b/x
plot(x, y, type = "l", main = "untransformed")
plot(1/x, y, type = "l", main = "1/x")
```

```{r}
# power
x <- seq(from = 1, to = 100, length.out = 100)
y <- a * x^b
plot(x, y, type = "l", main = "untransformed")
plot(x^b, y, type = "l", main = "x^b")
```

```{r}
# exp
x <- seq(from = 1, to = 10, length.out = 100)
y <- a * exp(b * x)
plot(x, y, type = "l", main = "untransformed")
plot(x, log(y), type = "l", main = "log(y)")
```

