---
title: "Mod6-7 (Wk 3)"
author: "Julianna D."
date: "2023-09-21"
output: html_document
---
# Module 6 Exploratory Data Analysis

## Challenge 1
```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(tidyverse)
```

What are the median area and population size of all countries in the dataset? HINT: There are a couple of ways to do this… try summary() and median() (for the latter, you’ll need to use the na.rm = TRUE argument)
Answer: Median area = 610; Median population size = 4.912e+06
```{r}
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/Country-Data-2016.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
summary(d)
names(d)
```

Create a new pop_density variable in your data frame which is population / area. What are the 10 most dense countries? The 10 least dense? HINT: check out the order() function
```{r}
d$density <- d$population/d$area
d <- d[order(-d$density), ]
d[1:10, ]
```

```{r}
d <- d[order(d$density), ]
d[1:10, ]
```

Extract data from all countries beginning with the letters “A” through “F”. What are the mean area and population size of these countries?
```{r}
new <- d[grep("^[A-F]", d$country), ]
summary(new)
```
Alternatively...
```{r}
mean(new$population, na.rm = TRUE)
mean(new$area, na.rm = TRUE)
```

## Challenge 2 Box & Bar Plots

```{r, warning=FALSE}
par(mfrow = c(2, 3))
boxplot(d$population)
boxplot(log(d$population))
boxplot(d$area)
boxplot(log(d$area))
barplot(d$population)
barplot(d$area)
```

## Challenge 3 Histograms
```{r, message=FALSE}
par(mfrow = c(1, 2))  # gives us two panels
attach(d)
hist(log(population), freq = FALSE, col = "red", main = "Plot 1", xlab = "log(population size)",
    ylab = "density", ylim = c(0, 0.2))
hist(log(area), freq = FALSE, col = "red", main = "Plot 2", xlab = "log(area)",
    ylab = "density", ylim = c(0, 0.2))
```
Density Plot
```{r}
par(mfrow = c(1, 1))  # set up one panel and redraw the log(population) histogram
hist(log(population), freq = FALSE, col = "white", main = "My Plot with Mean and Density",
    xlab = "log(population size)", ylab = "density", ylim = c(0, 0.2))
abline(v = mean(log(population), na.rm = TRUE), col = "blue")
lines(density(log(population), na.rm = TRUE), col = "green")
```
```{r}
detach(d)
```

## Challenge 4 Tables

```{r}
sort(table(d$govt_form), decreasing = TRUE)
```

## Challenge 5 For Multivariate Data
```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/KamilarAndCooperData.csv")
d <- read.csv(f, header = TRUE, stringsAsFactors = FALSE)
attach(d)
head(d)
```
```{r}
boxplot(log(Body_mass_female_mean) ~ Family, d)
detach(d)
```

```{r}
p <- ggplot(data = d, aes(x = Family, y = log(Body_mass_female_mean)))  #define the variables
p <- p + geom_boxplot()  #graph them in a boxplot
p <- p + theme(axis.text.x = element_text(angle = 90))  #put x-axis names at 90deg
p <- p + ylab("log(Female Body Mass)")  #rename y-axis title
p 
```

## Challenge 6 Scatterplots
```{r, message=FALSE}
attach(d)
par(mfrow = c(1, 2))
plot(x = Body_mass_female_mean, y = Brain_Size_Female_Mean)
plot(x = log(Body_mass_female_mean), y = log(Brain_Size_Female_Mean))
detach(d)
```
```{r}
p <- ggplot(data = d, aes(x = log(Body_mass_female_mean), y = log(Brain_Size_Female_Mean),
    color = factor(Family)))  # first, we build a plot object and color points by Family
p <- p + xlab("log(Female Body Mass)") + ylab("log(Female Brain Size)")  # then we modify the axis labels
p <- p + geom_point()  # then we make a scatterplot
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p
```
```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```

```{r}
p <- p + geom_smooth(method = "lm", fullrange = TRUE)
p
```
## Challenge 7
```{r}
p <- ggplot(data = d, aes(x = log(Body_mass_female_mean), y = log(MaxLongevity_m)))
p <- p + geom_point()
p <- p + geom_smooth(method = "lm")
p
```
```{r}
aggregate(d$Body_mass_female_mean ~ d$Family, FUN = "mean", na.rm = TRUE)
```

```{r}
s <- filter(d, Family == "Hominidae" & Mass_Dimorphism > 2)
head(s)
```
```{r}
s <- select(d, Family, Genus, Body_mass_male_mean)
head(s)
```
```{r}
s <- rename(d, Female_Mass = Body_mass_female_mean)
head(s$Female_Mass)
```
```{r}
s <- mutate(d, Binomial = paste(Genus, Species, sep = " "))
head(s$Binomial)
```

```{r}
s <- summarise(d, avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
    na.rm = TRUE))
s
```

```{r}
byFamily <- group_by(d, Family)
byFamily
s <- summarise(byFamily, avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
    na.rm = TRUE))
s
```
Piping
```{r}
s <-                                                             #to create dataframe "s"
  d %>%                                                          #take dataframe "d"
  group_by(Family) %>%                                           #Group it by Family
  summarise(avgF = mean(Body_mass_female_mean, na.rm=TRUE),      #And calculate mean male BM
            avgM = mean(Body_mass_male_mean, na.rm=TRUE))        #And mean female BM
s
```

## Challenge 8
```{r}
s <- d %>%
    mutate(Binomial = paste(Genus, Species, sep = " ")) %>%
    select(Binomial, Body_mass_female_mean, Body_mass_male_mean, Mass_Dimorphism) %>%
    group_by(Binomial) %>%
    summarise(avgF = mean(Body_mass_female_mean, na.rm = TRUE), avgM = mean(Body_mass_male_mean,
        na.rm = TRUE), avgBMD = mean(Mass_Dimorphism, na.rm = TRUE))
s
```
Acccording to Kamilar & Cooper’s (2013) dataset, what is the average male and female size, and body mass dimorphism of my two main study species (vervet monkeys, Chlorocebus pygerythrus; and woolly monkeys, Lagothrix lagotricha)? Which has a larger average female body mass? Which is more sexually dimorphic?
- See output above

Compare the body size of my two main study taxa at the Family level (i.e., Cercopithecidae vs. Atelidae) by plotting (using {ggplot2}) the body mass of males and females and sexual dimorphism. If you can, make the Cercopithecid boxes green, and the Atelid boxes purple.
-- Struggling with this

# Module 7 Central Tendency and Variance

## Challenge 1
```{r}
x <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 25, 50, 100, 200, 1000)
gm1 <- function(x) {
    prod(x)^(1/length(x))
}
gm1(x)

gm2 <- function(x) {
    exp(mean(log(x)))
}
gm2(x)
```

## Challenge 2
```{r}
ss1 <- function(x) {
    sum((x - mean(x))^2)
}
ss1(x)
# This is equivalence to...
ss2 <- function(x) {
    sum(x^2) - length(x) * mean(x)^2
}
ss2(x)
# A shortcut!
ss3 <- function(x) {
    sum(x^2) - (sum(x))^2/length(x)
}
ss3(x)
```

Population Variance
```{r}
pop_v <- function(x) {
    sum((x - mean(x))^2)/(length(x))
}
pop_v(x)
```

## Challenge 3
```{r}
sample_v <- function(x) {
    sum((x - mean(x))^2)/(length(x) - 1)
}
sample_v(x)
```

## Challenge 4
```{r}
SE1 <- function(x) {
    sqrt(sample_v(x)/length(x))
}
SE1(x)

#With the var() function
SE2 <- function(x) {
    sqrt(var(x)/length(x))
}
SE2(x)
```

```{r}
library(sciplot)
se(x)
```

```{r}
set.seed(1)
x <- rnorm(10000, 0, 1)
hist(x)
```

```{r}
x <- seq(from = -4, to = 4, by = 0.01)
plot(x, dnorm(x), cex = 0.4)
```

```{r}
plot(x, pnorm(x), cex = 0.4)
```

```{r}
x <- seq(from = 0, to = 1, by = 0.01)
plot(qnorm(x), x, cex = 0.4)
```

Confidence Intervals
```{r}
x <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)
m <- mean(x)
n <- length(x)
v <- var(x)
s <- sd(x)
e <- sqrt(v/n)
upper <- mean(x) + qnorm(0.975, mean = 0, sd = 1) * se(x)
lower <- mean(x) + qnorm(0.025, mean = 0, sd = 1) * se(x)  # or lower <- mean(x) - qnorm(0.975)*se(x)
ci <- c(lower, upper)
ci
```

Alternative Way to find CI
```{r}
upper <- m + qnorm(0.975, mean = 0, sd = 1) * e
lower <- m + qnorm(0.025, mean = 0, sd = 1) * e  # or lower <- m - qnorm(0.975)*e
ci <- c(lower, upper)
ci
```

Bootstrapping CI
```{r}
set <- NULL  # sets up a dummy variable to hold our 10000 simulations
n <- 15
for (i in 1:10000) {
    set[i] <- mean(sample(x, n, replace = TRUE))
}
quantile(set)
quantile(set, c(0.025, 0.975))
```

How does the CI calculated this way, by simulation, compare to that calculated based on assuming a normal distribution?

-- Answer: Bootstrapping repeatedly samples observations with replacement from the original dataset to estimate the distribution of a statistic. This does not assume normal distribution (?)

How does the width of the CI change with decreasing or increasing n (the number of observations drawn from your sample with replacement)? For example, if we set n at 5? At 50? At 500?

-- Answer: As the sample size increases, the width of the CI tends to decrease. 