---
title: "Mod9-10 (Wk 5)"
author: "Julianna D."
date: "2023-10-05"
output: html_document
---

# Module 9 Intro to Statistical Inference

Review of Standard Errors & CIs
```{r}
n <- 1000
mu <- 3.5
sigma <- 4
v <- rnorm(n, mu, sigma)
s <- sample(v, size = 30, replace = FALSE)
m <- mean(s)
m
```

```{r}
sd <- sd(s)
sd
sem <- sd(s)/sqrt(length(s))
sem
lower <- m - qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci <- c(lower, upper)
ci
```

The Central Limit Theorem
```{r}
lambda <- 14
n <- 10
pop_se <- sqrt(lambda/n)  # the estimated SE
pop_se
```

```{r}
x <- NULL
for (i in 1:1000) {
    x[i] <- mean(rpois(n = n, lambda = lambda))
}
hist(x, breaks = seq(from = lambda - 4 * sqrt(lambda)/sqrt(n), to = lambda +
    4 * sqrt(lambda)/sqrt(n), length.out = 20), probability = TRUE)
```

```{r}
sd <- sd(x)  # st dev of the sampling distribution
sd
```

```{r}
qqnorm(x)
qqline(x)
```

```{r}
#Sample size of 100
n <- 100
pop_se <- sqrt(lambda/n)  # the estimated SE
pop_se
```
```{r}
x <- NULL
for (i in 1:1000) {
    x[i] <- mean(rpois(n = n, lambda = lambda))
}
hist(x, breaks = seq(from = lambda - 4 * sqrt(lambda)/sqrt(n), to = lambda +
    4 * sqrt(lambda)/sqrt(n), length.out = 20), probability = TRUE)
```
```{r}
sd <- sd(x)
sd
```
```{r}
qqnorm(x)
qqline(x)
```
We can convert these distributions to standard normals by subtracting off the expected population mean (λ
) and dividing by the standard error of the mean (an estimate of the standard deviation of the sampling distribution) and then plotting a histogram of those values along with a normal curve.
```{r}
curve(dnorm(x, 0, 1), -4, 4, ylim = c(0, 0.8))
z <- (x - lambda)/pop_se
hist(z, breaks = seq(from = -4, to = 4, length.out = 20), probability = TRUE,
    add = TRUE)
```

```{r}
n <- 100
x <- NULL
for (i in 1:1000) {
    x[i] <- sum(rpois(n = n, lambda = lambda))
}
hist(x, breaks = seq(min(x), max(x), length.out = 20), probability = TRUE)
```

## Challenge
```{r}
n <- 1000
x <- 856
phat <- x/n  # our estimate of pi
phat
```

```{r}
n*phat
n*(1-phat)
pop_se <- sqrt((phat) * (1-phat)/n)
pop_se
```
```{r}
curve(dnorm(x, mean = phat, sd = pop_se), phat - 4 * pop_se, phat + 4 * pop_se)
upper <- phat + qnorm(0.975) * pop_se
lower <- phat - qnorm(0.975) * pop_se
ci <- c(lower, upper)
polygon(cbind(c(ci[1], seq(from = ci[1], to = ci[2], length.out = 1000), ci[2]),
    c(0, dnorm(seq(from = ci[1], to = ci[2], length.out = 1000), mean = phat,
        sd = pop_se), 0)), border = "black", col = "gray")
abline(v = ci)
abline(h = 0)
```

```{r}
mu <- 0
sigma <- 1
curve(dnorm(x, mu, 1), mu - 4 * sigma, mu + 4 * sigma, main = "Normal Curve=red\nStudent's t=blue",
    xlab = "x", ylab = "f(x)", col = "red", lwd = 3)
for (i in c(1, 2, 3, 4, 5, 10, 20, 100)) {
    curve(dt(x, df = i), mu - 4 * sigma, mu + 4 * sigma, main = "T Curve", xlab = "x",
        ylab = "f(x)", add = TRUE, col = "blue", lty = 5)
}
```

```{r}
n <- 1e+05
mu <- 3.5
sigma <- 4
x <- rnorm(n, mu, sigma)
sample_size <- 30
s <- sample(x, size = sample_size, replace = FALSE)
m <- mean(s)
m
```

```{r}
sd <- sd(s)
sd
sem <- sd(s)/sqrt(length(s))
sem
lower <- m - qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci_norm <- c(lower, upper)
ci_norm
```

Now, let’s look at the CIs calculated based using the t distribution for the same sample size. For sample size 30, the difference in the CIs is negligible.
```{r}
lower <- m - qt(1 - 0.05/2, df = sample_size - 1) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qt(1 - 0.05/2, df = sample_size - 1) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci_t <- c(lower, upper)
ci_t
```

Sample=5, CI based on t dist. is much wider
```{r}
sample_size <- 5
s <- sample(x, size = sample_size, replace = FALSE)
m <- mean(s)
m
sd <- sd(s)
sd
sem <- sd(s)/sqrt(length(s))
sem
lower <- m - qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci_norm <- c(lower, upper)
ci_norm
```
Compare to t distribution: 
```{r}
lower <- m - qt(1 - 0.05/2, df = sample_size - 1) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qt(1 - 0.05/2, df = sample_size - 1) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci_t <- c(lower, upper)
ci_t
```

# Module 10 Classical Hypothesis Testing

## Challenge 1 One Sample T-tests
```{r}
library(curl)
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall21/woolly-weights.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

```{r}
x <- d$weight
m <- mean(x)
s <- sd(x)
n <- length(x)
sem <- s/sqrt(n)
mu <- 7.2
t <- (m - mu)/sem
t
```

```{r}
alpha <- 0.05
crit <- qt(1 - alpha/2, df = n - 1)  # identify critical values
test <- t < -crit || t > crit  # boolean test as to whether t is larger than the critical value at either tail
test <- abs(t) > crit
t.test(x = x, mu = mu, alternative = "two.sided")
```

## Challenge 2 Two Sample T-test
```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/colobus-weights.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

```{r}
x <- d$weight[d$sex == "male"]
y <- d$weight[d$sex == "female"]
par(mfrow = c(1, 2))
boxplot(x, ylim = c(4.5, 8), main = "Weight (kg)", xlab = "Males")
boxplot(y, ylim = c(4.5, 8), main = "Weight (kg)", xlab = "Females")
```

```{r}
m1 <- mean(x)
m2 <- mean(y)
mu <- 0  # default argument
s1 <- sd(x)
s2 <- sd(y)
n1 <- length(x)
n2 <- length(y)
```

```{r}
t <- (m2 - m1 - mu)/sqrt(s2^2/n2 + s1^2/n1)
t
```

```{r}
alpha <- 0.05
crit <- qt(1 - alpha/2, df = n - 1)
crit
```

```{r}
test <- t < -crit || t > crit  # boolean test
test <- abs(t) > crit
test
```

```{r}
df <- (s2^2/n2 + s1^2/n1)^2/((s2^2/n2)^2/(n2 - 1) + (s1^2/n1)^2/(n1 - 1))
df
```

```{r}
t <- t.test(x = x, y = y, mu = 0, alternative = "two.sided")
t
```

Samples with Equal Variances
```{r}
s <- sqrt((((n1 - 1) * s1^2) + ((n2 - 1) * s2^2))/(n1 + n2 - 2))
t <- (m2 - m1 - mu)/(sqrt(s^2 * (1/n1 + 1/n2)))
t
df <- n1 + n2 - 2
df
```

```{r}
t <- t.test(x = x, y = y, mu = 0, var.equal = TRUE, alternative = "two.sided")
t
```

```{r}
var(x)/var(y)
```

```{r}
vt <- var.test(x, y)
vt
```

## Challenge 3

```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall23/iqs.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

```{r}
x <- d$IQ.before - d$IQ.after
m <- mean(x)
mu <- 0  # can leave this out
s <- sd(x)
n <- length(x)
sem <- s/sqrt(n)
par(mfrow = c(1, 2))
boxplot(d$IQ.before, ylim = c(115, 145), main = "IQ", xlab = "Before")
boxplot(d$IQ.after, ylim = c(115, 145), main = "IQ", xlab = "After")
```

```{r}
t <- (m - mu)/sem
t
alpha <- 0.05
crit <- qt(1 - alpha/2, df = n - 1)  # identify critical values
crit
test <- t < -crit || t > crit  # boolean test
test
t.test(x, df = n - 1, alternative = "two.sided")
```

Testing Sample Props One Sample Z Test

## Challenge 4
```{r}
v <- c(0, 1, 1, 1, 0, 1, 0, 0, 1, 1, 0, 0, 1, 1, 0, 1, 1, 0, 0, 1, 1, 1, 0,
    1, 1, 0, 1, 0, 1, 1)
```

```{r}
phat <- mean(v)
phat
```

```{r}
pi <- 0.8
n <- 30
z <- (phat - pi)/sqrt(pi * (1 - pi)/30)
z
```

```{r}
p <- pnorm(z, lower.tail = TRUE)
p
```

```{r}
lower <- phat - qnorm(0.975) * sqrt(phat * (1 - phat)/30)
upper <- phat + qnorm(0.975) * sqrt(phat * (1 - phat)/30)
ci <- c(lower, upper)
ci
```

```{r}
pt <- prop.test(x = sum(v), n = length(v), p = 0.8, conf.level = 0.95, correct = FALSE,
    alternative = "less")
pt
```

Comparing Sample Props 2 Sample Z Tests

## Challenge 5

```{r}
v1 <- c(1, 0, 1, 1, 0, 0, 1, 1, 0, 0, 1, 1, 1, 1, 0, 0, 1, 1, 0, 0, 1, 1, 0,
    1, 0)
v2 <- c(1, 1, 0, 0, 1, 1, 1, 1, 0, 1, 0, 1, 1, 1, 0, 1, 0, 1, 1, 1, 1, 1, 0,
    0, 1, 1, 0, 1, 1, 1)
```

```{r}
pstar <- (sum(v1) + sum(v2))/(length(v1) + length(v2))
pstar
```

```{r}
phat1 <- mean(v1)
phat1
phat2 <- mean(v2)
phat2
```

```{r}
pi <- 0
z <- (phat2 - phat1)/sqrt((pstar * (1 - pstar)) * (1/length(v1) + 1/length(v2)))
z
```

```{r}
p <- 1 - pnorm(z, lower.tail = TRUE) + pnorm(z, lower.tail = FALSE)
p
```

```{r}
crit <- qnorm(1 - alpha/2)  # identify critical values
crit
```

```{r}
test <- p < -crit || p > crit  # boolean test
test
```

```{r}
pt <- prop.test(x = c(sum(v2), sum(v1)), n = c(length(v2), length(v1)), alternative = "two.sided",
    correct = FALSE)
pt
```

